---
title: "In-class Exercise 4"
author: "Javier Lian"
---

## Installing and Loading the R Packages

```{r}
pacman::p_load(sf, raster, spatstat, sparr, tmap, tidyverse)
```

## Importing and Preparing Study Area

Importing study area

```{r}
#| eval: false
kbb <- st_read(
  dsn = 'data/rawdata',
  layer = "Kepulauan_Bangka_Belitung"
) %>%
  st_transform(crs = 32748)
```

```{r}
kbb_sf <- st_read(
  dsn = 'data/rawdata',
  layer = "Kepulauan_Bangka_Belitung"
) %>%
  st_union() %>%
  st_zm(drop = TRUE, what = "ZM") %>%
  st_transform(crs = 32748)
```

### Converting OWIN

```{r}
kbb_owin <- as.owin(kbb_sf)
kbb_owin
```

```{r}
class(kbb_owin)
```

## Importing and Preparing Forest Fire data

```{r}
fire <- read_csv("data/rawdata/forestfires.csv")
```

Revised code chunk to convert tibble dataframe into sf dataframe.

Removes longitude and latitude columns and insert a new column Geometry (lon, lat)

```{r}
fire_sf <- read_csv("data/rawdata/forestfires.csv") %>%
  st_as_sf(
    coords = c("longitude", "latitude"),
    crs = 4326
  ) %>%
  st_transform(crs = 32748)
```

```{r}
fire_sf <- fire_sf %>%
  mutate(DayOfYear = yday(acq_date)) %>%
  mutate(Month_num = month(acq_date)) %>%
  mutate(Month_fac = month(acq_date, label = TRUE, abbr = FALSE))
```

## Visualising the Fire Points

Overall plot

```{r}
tm_shape(kbb_sf) +
  tm_polygons() +
  tm_shape(fire_sf) +
  tm_dots()
```

Visualising by months

```{r}
tm_shape(kbb_sf) +
  tm_polygons() +
  tm_shape(fire_sf) +
  tm_dots(size = 0.1) +
  tm_facets(
    by = "Month_fac",
    free.coords = FALSE, #prevent zooming to coords, show the entire map
    drop.units = TRUE
  )
```

## Extracting forest fires by Month

```{r}
fire_month <- fire_sf %>%
  select(Month_num)
```

## Creating ppp

```{r}
fire_month_ppp <- as.ppp(fire_month)
fire_month_ppp
```

Check the output is in the correct object class

```{r}
summary(fire_month_ppp)
```

Check for duplicates

```{r}
any(duplicated(fire_month_ppp))
```

Including Owin object

```{r}
fire_month_owin <- fire_month_ppp[kbb_owin]
summary(fire_month_owin)
```

```{r}
plot(fire_month_owin)
```

Computing Spatio-temporal KDE

```{r}
st_kde <- spattemp.density(fire_month_owin)
summary(st_kde)

# h = 15102.45 tells us the distance in meters
# lambda 0.0304 tells us the month
```

Plotting the spatio-temporal KDE object

```{r, fig.width = 12, fig.height = 10}
  trims <- c(7, 8, 9, 10, 11, 12)
  par(mfcol = c(2, 3))
  
  for (i in trims) {
    plot(st_kde, i,
         override.par = FALSE,
         fix.range = TRUE,
         main = paste("KDE at month", i))
  }
```

Plotting in year

```{r}
fire_yday <- fire_sf %>%
  select(DayOfYear)
fire_yday_ppp <- as.ppp(fire_yday)
fire_yday_owin <- fire_yday_ppp[kbb_owin]
kde_yday <- spattemp.density(fire_yday_owin)
summary(kde_yday)
```
