---
title: "In-class Exercise 5"
author: "Javier Lian"
---

## Loading the package

```{r}
pacman::p_load(sf, spdep, tmap, tidyverse, knitr, GWmodel)
```

Preparing the data

```{r}
hunan_2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```

```{r}
#| eval: false
hunan_sf <- st_read(
  dsn = "data/geospatial",
  layer = "Hunan"
)
hunan_sf <- left_join(hunan_sf, hunan_2012) %>% 
  select(1, 3, 7, 15, 16, 31, 32)
```

```{r}
#| eval: false
write_rds(hunan_sf, "data/rds/hunan_sf.rds")
```

```{r}
#| echo: false
hunan_sf <- read_rds("data/rds/hunan_sf.rds")
```

### Converting to SpatialPolygonDataFrame

```{r}
hunan_sp <- hunan_sf %>% as_Spatial()
```

### Geographically Weighted Summary Statistics with adaptive bandwidth

Determine adaptive bandwidth

::: panel-tabset
### AIC

```{r}
bw_AIC <- bw.gwr(
  GDPPC ~1,
  data = hunan_sp,
  approach = "AIC",
  adaptive = TRUE,
  kernel = "bisquare",
  longlat = T
)
```

### Cross-Validation

```{r}
bw_AIC <- bw.gwr(
  GDPPC ~1,
  data = hunan_sp,
  approach = "CV",
  adaptive = TRUE,
  kernel = "bisquare",
  longlat = T
)
```
:::

### Geographically Weighted Summary Statistics with fixed bandwidth

Determine fixed bandwidth 

::: panel-tabset

### AIC

```{r}
bw_AIC <- bw.gwr(
  GDPPC ~1,
  data = hunan_sp,
  approach = "AIC",
  adaptive = FALSE,
  kernel = "bisquare",
  longlat = T
)
```

### Cross-Validation

```{r}
bw_AIC <- bw.gwr(
  GDPPC ~1,
  data = hunan_sp,
  approach = "CV",
  adaptive = FALSE,
  kernel = "bisquare",
  longlat = T
)
```
:::

### Geographically Weighted Summary Statistics with fixed bandwidth

Computing weighted summary statistics
```{r}
gwstat <- gwss(
  data = hunan_sp,
  vars = "GDPPC",
  bw = bw_AIC,
  adaptive = TRUE,
  kernel = "bisquare",
  longlat = T
)
```

Preparing the output data
```{r}
gwstat_df <- as.data.frame(gwstat$SDF)
```

Next, cbind() is used to append the newly derived data.frame onto hunan_sf sf dataframe
```{r}
hunan_gstat <- cbind(hunan_sf, gwstat_df)
```

```{r}
tm_shape(hunan_gstat) +
  tm_fill(
    "GDPPC_LM",
    n = 5,
    style = "quantile"
  ) +
  tm_borders(alpha = 0.5) +
  tm_layout(
    main.title = "Distribution of geographically weighted mean",
    main.title.position = "center",
    legend.text.size = 1.2,
    legend.height = 1.50,
    legend.width = 1.50,
    frame = TRUE
  )
```