---
title: "Hands-on Exercise 12"
author: "Javier Lian"
---

## Getting Started

## **Installing and Loading R packages**

```{r}
pacman::p_load(sf, spdep, GWmodel, SpatialML, 
               tmap, rsample, yardstick, tidyverse,
               knitr, kableExtra, spatialRF)
```

## **Preparing Data**

### **Reading data file to rds**

Reading the input data sets. It is in simple feature data frame.

```{r}
mdata <- read_rds("data/rds/mdata.rds")
```

### **Data Sampling**

The entire data are split into training and test data sets with 65% and 35% respectively by using *initial_split()* of **rsample** package. rsample is one of the package of tigymodels.

```{r}
set.seed(1234)
HDB_sample <- mdata %>%
  sample_n(1500)
```

### Checking of overlapping point

```{r}
overlapping_points <- HDB_sample %>%
  mutate(overlap = lengths(st_equals(., .)) > 1)
summary(overlapping_points$overlap)
```

### Spatial Jitter

```{r}
HDB_sample <- HDB_sample %>%
  st_jitter(amount = 5)
```

## Data Sampling

```{r}
set.seed(1234)
resale_split <- initial_split(HDB_sample, prop = 0.67/10,)

train_data <- training(resale_split)
test_data <- testing(resale_split)
```

```{r}
write_rds(train_data, "data/rds/train_data.rds")
write_rds(test_data, "data/rds/test_data.rds")
```

### Multicollinearity check

```{r, fig.width=10, fig.height=10}
mdata_nogeo <- mdata %>%
  st_drop_geometry()
ggstatsplot::ggcorrmat(mdata_nogeo[, 2:17])
```

### Predictive Modelling with GWR

```{r}
gwr_bw_train_ad <- bw.gwr(resale_price ~ floor_area_sqm +
                            storey_order + remaining_lease_mths +
                            PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                            PROX_MRT + PROX_PARK + PROX_MALL +
                            PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                            WITHIN_350M_CHILDCARE + WITHIN_350M_BUS + WITHIN_1KM_PRISCH,
                          data=train_data,
                          approach="CV",
                          kernel="gaussian",
                          adaptive=TRUE,
                          longlat=FALSE)
```

### Saving Predicted values

```{r}
gwr_pred <- gwr.predict(formula = resale_price ~
                          floor_area_sqm + storey_order +
                          remaining_lease_mths + PROX_CBD + 
                          PROX_ELDERLYCARE + PROX_HAWKER + 
                          PROX_MRT + PROX_PARK + PROX_MALL + 
                          PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                          WITHIN_350M_CHILDCARE + WITHIN_350M_BUS + 
                          WITHIN_1KM_PRISCH, 
                        data=train_data, 
                        predictdata = test_data, 
                        bw=20, 
                        kernel = 'gaussian', 
                        adaptive=TRUE, 
                        longlat = FALSE)
gwr_pred_df <- as.data.frame(
  gwr_pred$SDF$prediction
) %>%
  rename(gwr_pred = "gwr_pred$SDF$prediction")
gwr_pred_df
```

## Predictive Modelling with RF method

```{r}
coords <- st_coordinates(HDB_sample)
coords_train <- st_coordinates(train_data)
coords_test <- st_coordinates(test_data)

train_data_nogeom <- train_data %>%
  st_drop_geometry()
```

### Calibrating RF model

```{r}
set.seed(1234)
rf <- ranger(resale_price ~ floor_area_sqm +
                            storey_order + remaining_lease_mths +
                            PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                            PROX_MRT + PROX_PARK + PROX_MALL +
                            PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                            WITHIN_350M_CHILDCARE + WITHIN_350M_BUS + WITHIN_1KM_PRISCH, data=train_data_nogeom)
```

### Preparing Test Data

```{r}
test_data_nogeom <- cbind(
  test_data, coords_test) %>%
  st_drop_geometry()
```

### Predicting with rf

```{r}
rf_pred <- predict(rf, data = test_data_nogeom)
```

### Saving the predicted values

```{r}
rf_pred_df <- as.data.frame(rf_pred$predictions) %>%
  rename(rf_pred = "rf_pred$predictions")
```

### Predictive Modelling with SpatialML

```{r}
set.seed(1234)
grf_ad <- grf(resale_price ~ floor_area_sqm +
                            storey_order + remaining_lease_mths +
                            PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                            PROX_MRT + PROX_PARK + PROX_MALL +
                            PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                            WITHIN_350M_CHILDCARE + WITHIN_350M_BUS + WITHIN_1KM_PRISCH, 
              dframe = train_data_nogeom,
              bw=20,
              kernel="adaptive",
              coords=coords_train)
```

### Predicting the test data

```{r}
grf_pred <- predict.grf(grf_ad,
                        test_data_nogeom,
                        x.var.name="X",
                        y.var.name="Y",
                        local.w=1,
                        global.w=0)
```

### Saving the predicted values

```{r}
grf_pred_df <- as.data.frame(grf_pred)
```

## Model Comparison

```{r}
test_data_pred <- test_data %>%
  select(resale_price) %>%
  cbind(gwr_pred_df) %>%
  cbind(rf_pred_df) %>%
  cbind(grf_pred_df)
```

### Transposing Data

```{r}
test_longer <- test_data_pred %>%
  st_drop_geometry() %>%
  pivot_longer(cols = ends_with("pred"),
               names_to = "model",
               values_to = "predicted")
```

### Renaming

```{r}
model_labels <- c(
  gwr_pred = "gwr",
  rf_pred = "Random Forest",
  grf_pred = "gwRF"
)

test_longer <- test_longer %>%
  mutate(model = recode(
    model, !!!model_labels
  ))
```

## Computing RMSE

```{r}
rmse_results <- test_longer %>%
  group_by(model) %>%
  rmse(truth = resale_price,
       estimate = predicted) %>%
  rename(rmse = .estimate) %>%
  select(model, rmse)
```

## Model Comparison Plots

```{r}
ggplot(rmse_results,
       aes(x = reorder(model, rmse),
           y = rmse,
           fill = "skyblue")) +
  geom_bar(stat = "identity",
           fill = "skyblue",
           color = "black",
           width = 0.7) +
  labs(title = "RMSE Comparison of Mod",
       y = "RMSE",
       x = "Model") +
  theme_minimal()
```

### Scatter Plots

```{r}
test_longer <- test_longer %>%
  left_join(rmse_results,
            by = "model")

ggplot(data = test_longer,
       aes(x = predicted,
           y = resale_price)) +
  facet_wrap(~ model) +
  geom_point() +
  geom_text(data = test_longer,
            aes(x = Inf, y = Inf,
                label = paste("RMSE:", round(rmse, 2))),
            hjust = 1.1, vjust = 1.1,
            color = "black", size = 4)
```

## Variable Importance

### Preparing the output

```{r}
var_imp <- data.frame(
  Variable = names(grf_ad$Global.Model$variable.importance),
  Importance = grf_ad$Global.Model$variable.importance
)
```

### The plot

```{r}
ggplot(var_imp, aes(x = reorder(Variable,
                                Importance),
                    y = Importance)) +
  geom_bar(stat = "identity",
           fill = "skyblue") +
  coord_flip() +
  labs(
    title = "Variable Importance from Ranger Model",
    x = "Variables",
    y = "Importance"
  ) +
  theme_minimal()
```
