---
title: "In-class Exercise 2"
author: "Javier Lian"
---
## Working with Master Plan 2019 Subzone Boundary Data

::: panel-tabset
### Tasks

1.  Create a sub folder called data in In-class_Ex02 folder

2.  If necessary visit data.gov.sg and download "Master Plan 2024 Subzone Boundary (web)" from the portal. You are required to download the ESRI shapefile and kml file.

3.  Type the code chunk to import the "Master Plan 2024 Subzone Boundary (web)" shapefile and kml file, save them in sf simple features data frame.

### Code

::: \[style = "fontsize: 1.5em"\]

```{r}
pacman::p_load(sf, tidyverse)
```

This code chunk imports shape file.

```{r}
mpsz14_shp <- st_read(
  dsn = "data/geospatial",
  layer = "MP14_SUBZONE_WEB_PL"
)
```

This code chunk imports kml file (error).

```{r}
class(mpsz14_shp)
#| eval: false
mpsz14_kml <- st_read("data/aspatial/MasterPlan2014SubzoneBoundaryWebKML.kml")
```
## Method for Resolving MasterPlan2014 kml File Error
```{r}
st_write(
  mpsz14_shp,
  "data/aspatial/MasterPlan2014SubzoneBoundaryWebKML.kml",
  delete_dsn = TRUE
)
```
:::

## Working with Master Plan 2019 Subzone Boundary Data

::: panel-tabset
### Task

1.  Visit data.gov.sg and download Master Plan 2019 Subzone Boundary (No Sea) from the portal. You are required to download both kml files.

2.  Move MPSZ-2019 shapefile provided for In-Class Exercise 1 folder on elearn to data sub-folder of In-Class_Ex02.

3.  Write a code chunk to import Master Plan 2019 Subzone Boundary (No Sea) kml and MPSZ-2019 into sf simple feature data.frame.

### Import Shapefile

```{r}
mpsz19_shp <- st_read(
  dsn = "data/geospatial",
  layer = "MP14_SUBZONE_WEB_PL"
)
```

### Import kml

```{r}
mpsz19_kml <- st_read(
  "data/aspatial/MasterPlan2019SubzoneBoundaryNoSeaKML.kml"
)
```
:::

::: panel-tabset
### Import MPSZ-19 Shape
```{r}
mpsz19_shp <- st_read(
  dsn = "data/geospatial",
  layer = "MPSZ-2019"
) %>%
  st_transform(crs = 3414)
```
:::

## Working with Population Data
::: panel-tabset
### Task
1. Read the csv data into a variable popdata
### Code
```{r}
popdata <- read_csv("data/aspatial/respopagesextod2023.csv")
```
:::

## Data Wrangling
::: panel-tabset

### Task
1. Write a code chunk to prpare a data frame showing population by Planning Area and Planning Subzone.

### Code
```{r}
popdata2023 <- popdata %>%
  group_by(PA, SZ, AG) %>%
  summarise(`POP` = sum(`Pop`)) %>%
  ungroup() %>%
  pivot_wider(
    names_from = AG,
    values_from = POP
  )
```
:::